Probabilistic Programming

### 2014-04-21 paper/paper.tex
Start generate section

More formally, given a depth $d$ (indicating how many nodes we should backtrace) for a given
${node}_d$ and feature ${edge}_d$, the generation of ${node}_d$'s child node given by the
transition ${edge}_d$, can be expressed by the following function $f$,

\begin{align*}
f({node}_1, {edge}_1, ..., {node}_{d}, {edge}_{d}) = \begin{dcases*}
		{newnode_1},	&	$P({newnode_1})$ \\
		{newnode_2},	&	$P({newnode_2})$ \\
		...\\
		{newnode_n},    &   $P({newnode_n})$ \\
	\end{dcases*}
\end{align*}

such that ${Path}({node}_1, {edge}_1) = {node}_2$, ${Path}({node}_2, {edge}_2) = {node}_3$, ..., and ${Path}({node}_d, {edge}_d) = {newnode}_k$, where ${newnode}_k$ is randomly selected from ${newnode}_1, ..., {newnode}_k$ given the probabilities $P({newnode_1}), ..., P({newnode_n})$, and

\begin{align*}
\sum\limits_{i = 0}^n P({newnode_i}) = 1.
\end{align*}

From this it can be seen that generating programs is simply a repetitive application of this function, where for the chosen ${newnode}_k$, we can generate each child of ${newnode}_k$ with, $\forall {edge} \in {Features}({newnode}_k)$,

\begin{align*}
{Path}({newnode}_k, {edge}) = f({node}_2, {edge}_2, ..., {node}_{d}, {edge}_{d}, {newnode}_k, {edge}).
\end{align*}

@@

The data structure is a hash table (also commonly called a map, dictionary, or associative array), that, on a given input, returns another hash table for paths that begin with that input. This behavior is repeated until the number of inputs reaches the desired depth, at which point a final hash table, mapping the node type to its probability distribution is returned. This, in effect, is modularizing the function $f$ described above, so that, say for ${depth} = 2$,

\begin{align*}
lookup(lookup(lookup(lookup({hash}, {node}_1), {edge}_1), {node}_2), {edge}_2) = \textbf{P}({nodes}).
\end{align*}

@@

For each AST, we walk through every node in the tree by recursively exploring each feature\footnote{While we use \emph{follow} as if it were a feature edge, programs encoded in the AST's we parse use arrays of nodes (rather than linked-lists as we represent them) for a collection of statements. The given pseudocode demonstrates how we can transform an array into a linear sequence of edges.} of each node. In each call to explore a node's feature, we maintain the path up until that node by appending to it the node type and the name of the feature to reach the wanted child. We trim the beginning of the path off so that the length of the path is appropriate with respect to the chosen depth (i.e. so that $|path| = 2*d$)\footnote{Note that this requires that the path is at least of length $2*depth$. If the path is shorter than this, it represents not having at least $d$ levels of context. This can occur, say, for the first statement in the entire program, where there are no preceding nodes in its AST path. For these cases, we add the ``{\tt \_null}'' feature until we reach the desired depth (omitted in the pseudocode.)}, create that path through the layers of hash tables, if it doesn't already exist, add the node type to the probability distribution if it isn't already, and then increment a counter associated with that node type. We also increment a ``{\tt \_total}'' counter for the convenience of the generator program.

@@

>>>>

More formally, given a depth $d$ (indicating how many nodes we should backtrace) for a given
${node}_d$ and feature ${edge}_d$, the generation of ${node}_d$'s child node given by the
transition ${edge}_d$, can be expressed by the following function $f$, \begin{align*}
f({node}_1, {edge}_1, ..., {node}_{d}, {edge}_{d}) = \begin{dcases*}
		{newnode_1},	&	$P({newnode_1})$ \\
		{newnode_2},	&	$P({newnode_2})$ \\
		...\\
		{newnode_n},    &   $P({newnode_n})$ \\
	\end{dcases*}
\end{align*}

such that ${Path}({node}_1, {edge}_1) = {node}_2$, ${Path}({node}_2, {edge}_2) = {node}_3$, ..., and ${Path}({node}_d, {edge}_d) = {newnode}_k$, where ${newnode}_k$ is randomly selected from ${newnode}_1, ..., {newnode}_k$ given the probabilities $P({newnode_1}), ..., P({newnode_n})$, and \begin{align*}
\sum\limits_{i = 0}^n P({newnode_i}) = 1.
\end{align*}

From this it can be seen that generating programs is simply a repetitive application of this function, where for the chosen ${newnode}_k$, we can generate each child of ${newnode}_k$ with, $\forall {edge} \in {Features}({newnode}_k)$, \begin{align*}
{Path}({newnode}_k, {edge}) = f({node}_2, {edge}_2, ..., {node}_{d}, {edge}_{d}, {newnode}_k, {edge}).
\end{align*}

@@

The data structure is a hash table (also commonly called a map, dictionary, or associative array), that, on a given input, returns another hash table for paths that begin with that input. This behavior is repeated until the number of inputs reaches the desired depth, at which point a final hash table, mapping the node type to its probability distribution is returned. This, in effect, is modularizing the function $f$ described above, so that we can reach the probability distribution with minimal code and data duplication. This is demonstrated by following example, for  ${depth} = 2$. \begin{align*}
lookup(lookup(lookup(lookup({model}, {node}_1), {edge}_1), {node}_2), {edge}_2) = \textbf{P}({nodes}).
\end{align*}

@@

For each AST, we walk through every node in the tree by recursively exploring each feature\footnote{While we use \emph{follow} as if it were a feature edge, programs encoded in the AST's we parse use arrays of nodes (rather than linked-lists as we represent them) for a collection of statements. The given pseudocode demonstrates how we can transform an array into a linear sequence of edges.} of each node. In each call to explore a node's feature, we maintain the path up until that node by appending to it the node type and the name of the feature to reach the wanted child. We trim the beginning of the path off so that the length of the path is appropriate with respect to the chosen depth (i.e. so that $|path| = 2*d$).\footnote{Note that this requires that the path is at least of length $2*depth$. If the path is shorter than this, it represents not having at least $d$ levels of context. This can occur, say, for the first statement in the entire program, where there are no preceding nodes in its AST path. For these cases, we add the ``{\tt \_null}'' feature until we reach the desired depth (omitted in the pseudocode.)} We traverse through the layers of hash tables, creating the path elements if they don't already exist, add the node type to the probability distribution if it isn't already, and then increment a counter associated with that node type. We also increment a ``{\tt \_total}'' counter for the convenience of the generator program.

@@

Random programs can be generated rather easily using the approach explained in Section 2.2. The program tree is built up one node at a time, each one generated based on its unique AST path. We always begin a program with a ``Program'' node
